A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-16 papers/static-checker/ecoop2024/ecoop2024.tex
add related work section

%TODO mandatory: add short abstract of the document
\begin{abstract}
Formal verification is a requirement for all but the simplest smart
contracts on the blockchain. We developed XXX, a tool that is geared towards contracts
implemented in Michelson, the smart contract language of the Tezos
blockchain. It relies on symbolic execution and an SMT solver to
derive verification conditions that can be checked against the user's
specification. Specifications are written in a domain-specific
language tailored to Michelson's execution model.

We report the foundations of the design and implementation of XXX and
conduct two case studies that demonstrate the expressiveness of our tool.
\end{abstract}

@@

Formal verification can guarantee the correctness of smart contracts.
To this end we have developed XXX, a formal
verification tool for smart contracts on the Tezos blockchain, written in
Michelson, although our design is widely applicable. XXX relies on
symbolic execution backed by an SMT solver to simulate the
implementation of the smart contract language, helping to detect 
subtle errors. In addition, XXX comes with a domain-specific language that
allows users to precisely specify contract properties.
% By interacting with an SMT solver, it can handle a wide range of properties.
It streamlines the process of reviewing requirements,
uncovering hidden errors, and validating user-defined properties.
XXX provides a comprehensive solution to
mitigate potential pitfalls in blockchain-based applications.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                  Pair (highest_bidder: address)
                       (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance - Amount
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance - Amount, c) ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The actual specification consists of a precondition and a
postcondition with the usual partial correctness semantics. The
postcondition refers to components of the final storage by prefixing
\lstinline|post_| to their name. The variables \lstinline|Balance|,
\lstinline|Amount|, and \lstinline|Sender| correspond to the values
returned by the like-named Michelson instructions that return the
current balance of the contract, the amount transferred with the call,
and the address of the caller (sender).

@@

Our contributions are as follows.
\begin{enumerate}
\item Design and implementation of the smart contract specification
  language XXX.
\item Design and implementation of a symbolic interpreter for a
  significant subset of Michelson.
\item Two case studies where we apply XXX to verify real life smart
  contracts deployed on the Tezos blockchain:
  \begin{itemize}
  \item USDtz, an implementation of the FA~1.2 standard for financial
    applications;
  \item the Kolibri oracle contract that injects currency exchange
    rates from an external system.
  \end{itemize}
\end{enumerate}

@@

\atha{Add citations/references to the high-level languages.}
While contracts may be implemented directly in Michelson, it can also
be used as a compilation target. Indeed, several high-level languages
such as LIGO, SmartPy, and Liquidity have been created. These
languages provide a more readable syntax, resembling popular
programming languages, allowing developers to write Tezos smart
contracts more comfortably. Our tools are applicable to all of them
because they compile to Michelson.
\lstset{language=Caml}

@@

One of the main challenges with formal verification is that users are
often required to express  system properties in a specialized
specification language. This requirement poses two problems: it
demands expertise in formal verification from the specifier and
introduces a gap between the actual code and its specification. In our
framework, users are not obliged to possess formal verification
expertise. A basic understanding of Michelson data types and their
usage when interacting with a smart contract on the Tezos blockchain
is sufficient.
\begin{figure}[tp]
\begin{align*}
\textbf{cpt} &::= 
   \text{unit} 
   \Mid\ \text{never} 
   \Mid\ \text{bool} 
   \Mid\ \text{int}
   \Mid\ \text{nat}
   \Mid\ \text{string}
   \Mid\ \text{chain-id}
   \Mid\ \text{bytes}
   \Mid\ \text{mutez} \\
   &\Mid\ \text{key-hash}
   \Mid\ \text{key}
   \Mid\ \text{signature}
   \Mid\ \text{timestamp}
   \Mid\ \text{address} \\
   &\Mid\ \text{option}\ \textbf{cpt}
   \Mid\ \text{or}\ \textbf{cpt}\ \textbf{cpt}
   \Mid\ \text{pair}\ \textbf{cpt}\ \textbf{cpt} \\
\text{T, U} &::= 
    \text{cpt}
   \Mid\ \text{option}\ \text{T}
   \Mid\ \text{list}\ \text{T}
   \Mid\ \text{set}\ \textbf{cpt} 
   \Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \text{T}
   \Mid\ \text{ticket}\ \textbf{cpt} 
   \Mid\ \text{pair}\ \text{T}\ \text{U}
   \Mid\ \text{or}\ \text{T}\ \text{U}
   \Mid\ \text{lambda}\ \text{T}\ \text{U} \\
   &\Mid\ \text{map}\ \textbf{cpt}\ \text{T}
   \Mid\ \text{big-map}\ \textbf{cpt}\ \text{T}
   \Mid\ \text{chest} 
   \Mid\ \text{chest-key}
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}

\begin{figure}[tp]
\begin{align*}
\textbf{c} & ::= 
    \text{Balance}
   \Mid \text{post\_Balance}
   \Mid \text{Amount}
   \Mid \text{Sender}
   \Mid \text{Source} 
   \Mid \text{Now}
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id}
   \Mid \text{Self\_address}
   \Mid \text{Total\_voting\_power}
   \Mid \text{Voting\_power} \\
\textbf{t} &::= 
   \textbf{n}: \text{nat}
   \Mid\ \textbf{i}: \text{int}
   \Mid\  \textbf{s}: \text{string} 
   \Mid\  \textbf{b}: \text{bytes}
   \Mid\  \textbf{c}  
   \Mid\  \text{x} 
   \Mid\ \text{x}: \text{T} \\
   &\Mid\ \UNIT 
   \Mid\ \NEVER 
   \Mid\ \TRUE 
   \Mid\ \FALSE 
   \Mid\ \PAIR\ \text{t1 t2} \\
   &\Mid\ \LEFT\ \text{t}\  \text{T}
   \Mid\ \RIGHT\ \text{T}\ \text{t}
   \Mid\ \SOME\ \text{t}
   \Mid\ \NONE\ \text{T} 
   \Mid\ \text{\{t ; ... \}}: \text{list T}
   \Mid\ \text{\{t ; ... \}}: \text{set T} \\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{map T U}
   \Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{big\_map T U} \\
\textbf{e} & :: = \textbf{t}  \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ / } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e}, \textbf{e}, \textbf{e}) 
 \Mid \text{get\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{get\_big\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{check\_mem} (\textbf{e}, \textbf{e})  \\
& \Mid \text{get\_contract} (\textbf{e}, \textbf{e})  
 \Mid \text{contain} (\textbf{e}, \textbf{e})  
 \Mid \text{get\_substring} (\textbf{e}, \textbf{e}, \textbf{e})  
 \Mid \text{check\_sig} (\textbf{e}, \textbf{e}, \textbf{e})  \\
& \Mid \text{abs} (\textbf{e})  
 \Mid \text{is\_nat} (\textbf{e})  
 \Mid \text{to\_int} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{cons} (\textbf{e}, \textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{ $\wedge$ } \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e}
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
&\Mid \text{forall} (\textbf{e}, \textbf{p})
\Mid \text{exist} (\textbf{e}, \textbf{p})
\Mid \textbf{p} \text{ and } \textbf{p}\Mid \textbf{p} \text{ or } \textbf{p}  \Mid \textbf{p} \text{ nor } \textbf{p} \Mid  \text{ not } \textbf{p} \Mid \textbf{p} \text{ imply } \textbf{p} \\
\end{align*}
\caption{Syntax of terms, t and predicates, p}
\label{fig:syntax-terms-predicates}
\end{figure}

@@

\subsection{Simple contracts}
\label{sec:simple-contracts}

Listing~\ref{lst:smart-contract-property-specification} shows the
syntax of our property specification language for a simple contract. Here, \KMCONTRACT,
\KSPEC, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are
keywords. The \KNAME\ is replaced by the contract name. In the input
and output fields, the input and output terms are specified. 
\atha{What terms are supported? terms are not explained!}
Since
Michelson operates on a stack, we assume that the stack has only one
element, as specified in the input term. Similarly, the output
stack is assumed to be a stack with only one element, which is
specified in the output field. Thus, the syntax allows us to specify
incomplete contracts, too.
\pt{I disagree (though it would be a useful feature). To this end, there should be terms to cover stacks
  with more than one element. As it stands, the input term can only
speak about the top of the stack and the output term only about the
second component of the top of the stack, which is assumed to be a
pair.\\
Moreover, a specification might want to say something about the list
of operations, too, e.g., that tokens are transferred to some address.}

For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code.  For verification purposes we elide the operation list and represent the output solely by the term of the value in the storage after the execution concludes. In some cases, the output term can be omitted
for a complete contract, given that the input term must be a pair
of parameters and storage. Hence the storage term is part of the
input term. In such cases, the prefix \lstinline/post_/ is used
before the variables from the input term to indicate the corresponding
storage value after the execution.

The input and output terms represent Michelson terms, which can be
either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its
corresponding type declaration. Subsequent mentions can omit the type
declaration.  Formally, the pre- and postcondition instances are
formulas in first-order logic that can be handled by an SMT solver (in
our case, Z3). The specification language is constituted of predicates.  The syntax terms and  predicates are defined in Figure~\ref{fig:syntax-terms-predicates}. Formulas that are used in assertions or
added to solvers are terms of sort Boolean. Otherwise, terms of
Boolean and non-Boolean sort may be mixed in any combination where
sorts match up. Universal and existential quantifiers bind variables
to the scope of the quantified formula.
\atha{More explanation is needed about expressions and formulas. Why
  do you support left and right, but not lists? Why Never? What is the
meaning of transfer\_token, get\_map, get\_contract? The constants
must be explained (though you could do that in
Section~\ref{sec:background}). (Less common items could be omitted!)}

@@

Let us consider the contract \texttt{Sub}  in Listing
\ref{lst:sub-contract-specification} as an example of a property
specification.  Here, the program takes as input a stack with only one
element, which is a pair of two integers, symbolically specified in
the \lstinline|input| field. The program first duplicates the pair. It then unpairs the input pair to obtain two elements as two integers \lstinline/x/ and \lstinline/y/.
Next, the program performs a comparison between
these two integers. If \lstinline/x/ is greater than or equal to
\lstinline/y/, the program constructs a new stack with a solitary
element at the top representing the subtract of  \lstinline/x/ and
\lstinline/y/, designated as \lstinline/z/ by the \lstinline|output| directive. Otherwise, the execution
terminates with an error message containing the content
\lstinline|'Unexpected pair'|.
\atha{This Michelson code with its explanation might be an example for Section~\ref{sec:background}}
The specification continues by defining
pre- and postconditions, where the precondition states the assumption
that the integer \lstinline/x/ is greater than or equal to  \lstinline/y/. The postcondition asserts that the result
\lstinline/z/ is greater than or equal to the substract of \lstinline/x/ and \lstinline/y/, which is greater or equal to zero. The symbol \lstinline/&&/ stands for logical conjunction.
\atha{This specification is meaningless. Why not use $x\ge y$ as
  precondition and $z = x+y$ as postcondition?}

\subsection{More than one entrypoint}
\label{sec:more-than-one}

In XXX, we can specify several entrypoints at the same
time. This design enables us to specify and verify all entrypoints
at once and also to describe and verify relationships between
entrypoints.  There are two options of the syntax as shown in Listing~\ref {lst:multiple-entrypoint-specification-1} and Listing~\ref {lst:multiple-entrypoint-specification-2},
where the entrypoint name is written after the \lstinline/%/ symbol
and is followed by a property specification for the entrypoint. The
specification finishes with specifying entrypoint relations,
which we explain next.

@@

\atha{can we put the source code of the contract somewhere, e.g. in an appendix?}
To illustrate entrypoint relations, let us consider a smart contract
that models an online auction. The contract has two entrypoints,
\lstinline|%bid| and \lstinline|%close|, which serve as functions for
bidding and closing the contract, respectively.  The smart contract
storage is a pair containing as its first element a boolean value
\lstinline|auction_open| that indicates whether the contract is still
open for bidding or already closed. The second element is another
nested pair, where the first element stores the highest bidder's
address, and the second element is the contract owner's address. The
current highest bid is reflected in the balance of the contract.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                   Pair (highest_bidder: address)
                        (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance - Amount
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance - Amount, c);

  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open 
       && Sender = contract_owner  
       && get_contract (unit, contract_owner) = 
          Some (c: contract unit)              
    post-condition := not post_auction_open 
       && post_Balance = 0 
       && transfer_token (Unit, Balance, c)

  (%create -> %bid) with (auction_open = true)  
       and (Amount > Balance) 
       and get_contract (unit, highest_bidder) = 
           Some (c: contract unit)
  | (%create  -> %close) with (auction_open = true) 
       and (Sender = contract_owner)
       and get_contract (unit, contract_owner) = 
          Some (c: contract unit)
  | (%bid -> %bid) with (Amount > Balance - Amount) 
       and get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
  | (%bid -> %close) with (Sender = contract_owner) 
       and get_contract (unit, contract_owner) = 
            Some (c: contract unit)
  | not (%close -> %bid)
  | not (%close -> %close)
\end{lstlisting}

@@

Here is the first property that we want to check: if a user calls the
\lstinline/%bid/ entrypoint, when the contract is still open for
bidding and the bid (i.e., the amount transferred with the call) is
higher than the current bid (as specified in the
\lstinline/pre-condition/ field from line 10 to line 11 in Listing
\ref{lst:auction-contract-specification}), then the bid should
succeed.  In this case, the highest bidder has to be updated to the
sender (line 12 in the \lstinline/post-condition/ field), which is the
address of the caller, and the balance of the contract should be the
amount sent along with the call (line 13). Moreover, the previous
highest bidder should receive their bid, which means the auction
contract should transfer the \lstinline/Balance/ to the previous
bidder, \lstinline/highest_bidder/ (line 14). 
The predicate \lstinline/Transfer_token (data, amount, address)/
represents a smart contract call with three arguments: the data sent
along with the call, the amount of tokesn, and the address of the
receiver.  When the call is just a token transfer, the data field has
the value \lstinline/Unit/ of type \lstinline/unit/, indicating that
this is just a token transfer with no additional data. Another
important property is that the auction owner's address should not be
changed. That is, after calling the \lstinline/bid/ entrypoint, the
\lstinline/contract_owner/ should remain the same as before (line
15). The last property guarantees that when the contract owner (line
25) closes the open contract (line 24), the contract is first closed
(line 26), and then money (\lstinline/Balance/) should be sent to the
owner. Hence, the final auction balance is zero (lines 28 and 29).

\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}

>>>>

%TODO mandatory: add short abstract of the document
\begin{abstract}
Formal verification is a requirement for all but the simplest smart
contracts on the blockchain. We developed SCV, a tool that is geared towards contracts
implemented in Michelson, the smart contract language of the Tezos
blockchain. It relies on symbolic execution and an SMT solver to
derive verification conditions that can be checked against the user's
specification. Specifications are written in a domain-specific
language tailored to Michelson's execution model.

We report the foundations of the design and implementation of SCV and
conduct two case studies that demonstrate the expressiveness of our tool.
\end{abstract}

@@

Formal verification can guarantee the correctness of smart contracts.
To this end we have developed SCV, a formal
verification tool for smart contracts on the Tezos blockchain \cite{tezos-whitepaper}, written in
Michelson \cite{michelson}, although our design is widely applicable. SCV relies on
symbolic execution backed by an SMT solver to simulate the
implementation of the smart contract language, helping to detect 
subtle errors. In addition, SCV comes with a domain-specific language that
allows users to precisely specify contract properties.
% By interacting with an SMT solver, it can handle a wide range of properties.
It streamlines the process of reviewing requirements,
uncovering hidden errors, and validating user-defined properties.
SCV provides a comprehensive solution to
mitigate potential pitfalls in blockchain-based applications.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                  Pair (highest_bidder: address)
                       (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > pre_Balance 
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, Old_Balance, c) ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The actual specification consists of a precondition and a
postcondition with the usual partial correctness semantics. The
postcondition refers to components of the final storage by prefixing
\lstinline|post_| to their name. The variables
\lstinline|Amount|, and \lstinline|Sender| correspond to the values
returned by the like-named Michelson instructions that return, the amount transferred with the call, and the address of the caller (sender). The variables \lstinline|pre_Balance| and \lstinline|post_Balance| represent the previous balance of the contract right before and after the call, respectively. The current balance of the contract returned by the instruction \BALANCE\ includes \lstinline|pre_Balance| and the \lstinline|Amount|.

@@

Our contributions are as follows.
\begin{enumerate}
\item Design and implementation of the smart contract specification
  language SCV.
\item Design and implementation of a symbolic interpreter for a
  significant subset of Michelson.
\item Two case studies where we apply SCV to verify real life smart
  contracts deployed on the Tezos blockchain:
  \begin{itemize}
  \item USDtz, an implementation of the FA~1.2 standard for financial
    applications;
  \item the Kolibri oracle contract that injects currency exchange
    rates from an external system.
  \end{itemize}
\end{enumerate}

@@

While contracts may be implemented directly in Michelson, it can also
be used as a compilation target. Indeed, several high-level languages
such as LIGO \cite{ligo}, SmartPy \cite{smartpy}, and Liquidity \cite{liquidity} have been created. These languages provide a more readable syntax, resembling popular
programming languages, allowing developers to write Tezos smart
contracts more comfortably. Our tools are applicable to all of them
because they compile to Michelson.
\lstset{language=Caml}

@@

One of the main challenges with formal verification is that users are
often required to express  system properties in a specialized
specification language. This requirement poses two problems: it
demands expertise in formal verification from the specifier and
introduces a gap between the actual code and its specification. In our
framework, users are not obliged to possess formal verification
expertise. A basic understanding of Michelson data types and their
usage when interacting with a smart contract on the Tezos blockchain
is sufficient.
\begin{figure}[tp]
\begin{align*}
\textbf{cpt} &::= 
   \text{unit} 
   \Mid\ \text{never} 
   \Mid\ \text{bool} 
   \Mid\ \text{int}
   \Mid\ \text{nat}
   \Mid\ \text{string}
   \Mid\ \text{chain-id} \\
   & \Mid\ \text{bytes}
   \Mid\ \text{mutez} 
   \Mid\ \text{key-hash}
   \Mid\ \text{key}
   \Mid\ \text{signature}
   \Mid\ \text{timestamp} \\
   & \Mid\ \text{address} 
   \Mid\ \text{option}\ \textbf{cpt}
   \Mid\ \text{or}\ \textbf{cpt}\ \textbf{cpt}
   \Mid\ \text{pair}\ \textbf{cpt}\ \textbf{cpt} \\
\text{T, U} &::= 
    \text{cpt}
   \Mid\ \text{option}\ \text{T}
   \Mid\ \text{list}\ \text{T}
   \Mid\ \text{set}\ \textbf{cpt} 
   \Mid\ \text{operation} 
   \Mid\ \text{contract}\ \text{T} \\
  & \Mid\ \text{pair}\ \text{T}\ \text{U}
   \Mid\ \text{or}\ \text{T}\ \text{U}
   \Mid\ \text{lambda}\ \text{T}\ \text{U} 
   \Mid\ \text{map}\ \textbf{cpt}\ \text{T}
   \Mid\ \text{big-map}\ \textbf{cpt}\ \text{T}
\end{align*}
\caption{Types}
\label{fig:types}
\end{figure}

\begin{figure}[tp]
\begin{align*}
\textbf{c} & ::= 
    \text{pre\_Balance}
   \Mid \text{post\_Balance}
   \Mid \text{Amount}
   \Mid \text{Sender}
   \Mid \text{Source} 
   \Mid \text{Now}
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id}
   \Mid \text{Self\_address}\\
\textbf{t} &::= 
   \textbf{n}: \text{nat}
   \Mid\ \textbf{i}: \text{int}
   \Mid\  \textbf{s}: \text{string} 
   \Mid\  \textbf{b}: \text{bytes}
   \Mid\  \textbf{c}  
   \Mid\  \text{x} 
   \Mid\ \text{x}: \text{T} \\
   &\Mid\ \UNIT 
   \Mid\ \NEVER 
   \Mid\ \TRUE 
   \Mid\ \FALSE 
   \Mid\ \PAIR\ \text{t1 t2} \\
   &\Mid\ \LEFT\ \text{t}\  \text{T}
   \Mid\ \RIGHT\ \text{T}\ \text{t}
   \Mid\ \SOME\ \text{t}
   \Mid\ \NONE\ \text{T} 
   \Mid\ \text{\{t ; ... \}}: \text{list T}
   \Mid\ \text{\{t ; ... \}}: \text{set T} \\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{map T U}
   \Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{big\_map T U} \\
\textbf{e} & :: = \textbf{t}  \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ / } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& 
 \Mid \text{get\_map} (\textbf{e1}, \textbf{e2}) 
 \Mid \text{get\_big\_map} (\textbf{e1}, \textbf{e2})  \\
 &\Mid \text{get\_contract} (\textbf{e1}, \textbf{e2})   
 \Mid \text{get\_substring} (\textbf{e1}, \textbf{e2}, \textbf{e3})   \\
& \Mid \text{abs} (\textbf{e})  
 \Mid \text{to\_int} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{cons} (\textbf{e}, \textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{ $\wedge$ } \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e}
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e1}, \textbf{e2}, \textbf{e3})  
 \Mid \text{is\_nat} (\textbf{e})  
 \Mid \text{contain} (\textbf{e1}, \textbf{e2}) \\
& \Mid \textbf{p} \text{ and } \textbf{p}\Mid \textbf{p} \text{ or } \textbf{p}  \Mid \textbf{p} \text{ nor } \textbf{p} \Mid  \text{ not } \textbf{p} \Mid \textbf{p} \text{ => } \textbf{p} \\
\end{align*}
\caption{Syntax of terms, t and predicates, p}
\label{fig:syntax-terms-predicates}
\end{figure}

@@

\subsection{Simple contracts}
\label{sec:simple-contracts}
Listing~\ref{lst:smart-contract-property-specification} shows the
syntax of our property specification language for a simple contract. Here, \KMCONTRACT,
\KSPEC, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are
keywords. The \KNAME\ is replaced by the contract name. In the input
and output fields, the input and output terms are specified. The input and output terms represent Michelson terms, which can be either concrete values or symbolic values, including variables. The terms are specified using the syntax defined in Fig. \ref{fig:syntax-terms-predicates}. A term \textbf{t} comprises constants of the sort \text{nat} (natural number), \text{int} (integer), \text{string} (string), and \text{b} (bytes) along with their respective types. As a syntactic sugar, the type can be omitted except for the type \text{nat} to distinguish it from integer numbers. It also encompasses all blockchain constants \textbf{c}, such as Amount and Sender. When a variable appears for the first time, it must be accompanied by its corresponding type declaration. Subsequent mentions can omit the type declaration. The language supports essential Michelson types, as shown in Fig. \ref{fig:types}, where \textbf{cpt} denotes comparable types. The types include basic data types such as integer, string, list, set, pair, option and map. Additionally, there are some typical types related to the blockchain, such as \text{mutez} and \text{contract T}, representing types for tokens and smart contracts on the Tezos blockchain, respectively.

Since Michelson operates on a stack, we assume that the stack has only one
element, as specified in the input term. Similarly, the output
stack is assumed to be a stack with only one element, which is
specified in the output field. The syntax allows us to specify both
incomplete and complete smart contracts\footnote{The tool has been adapted to accommodate incomplete contract codes that take input stacks and produce output stacks with more than one element. This modification is feasible from a technical standpoint, as the current tool can effectively handle such cases.}. For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code.  For verification purposes we elide the operation list and represent the output solely by the term of the value in the storage after the execution concludes. In some cases, the output term can be omitted for a complete contract, given that the input term must be a pair
of parameters and storage. Hence the storage term is part of the
input term. In such cases, the prefix \lstinline/post_/ is used
before the variables from the input term to indicate the corresponding
storage value after the execution.

Formally, the pre- and postcondition instances are formulas in first-order logic that can be handled by an SMT solver (in our case, Z3). The specification language is constituted of predicates.  The syntax predicates are defined in Figure~\ref{fig:syntax-terms-predicates}. Formulas that are used in assertions or added to solvers are terms of sort Boolean. Otherwise, terms of
Boolean and non-Boolean sort may be mixed in any combination where
sorts match up. Several functions are defined, such as the function \text{get\_map (e1, e2)} for accessing an element in a map. This function returns \text{None} if there is no key \text{e1} in the map \text{e1}, or the corresponding value of the key otherwise. Another function, \text{get\_contract (e1, e2)}, casts the address \text{e2} to a real contract on the Tezos blockchain. It returns \text{None} if there is no existing contract on the Tezos block chain with an entrypoint of type \text{e1} associated with the address \text{e2}, or it returns the contract otherwise.

Some other functions are defined as predicates, such as \text{transfer\_token} (\text{e1}, \text{e2}, \text{e3}). This implies that there is an operation in the operation list, after executing the smart contract code, which indicates the transfer of \text{e2} amount of tokens to the address \text{e3} along with the data \text{e1}\footnote{While some additional functions and both universal and existential quantifiers are implemented, they pose no technical problems and are similar to existing supported functions. However, the conversion to Z3 formulas in the Z3 OCaml library takes some time.}.

@@

Let us consider a fragment code of the contract \texttt{Sub}  in Listing
\ref{lst:sub-contract-specification} as an example of a property
specification.  Here, the program takes as input a stack with only one
element, which is a pair of two integers, symbolically specified in
the \lstinline|input| field. The program first duplicates the pair. It then unpairs the input pair to obtain two elements as two integers \lstinline/x/ and \lstinline/y/.
Next, the program performs a comparison between
these two integers. If \lstinline/x/ is greater than or equal to
\lstinline/y/, the program constructs a new stack with a solitary
element at the top representing the subtract of  \lstinline/x/ and
\lstinline/y/, designated as \lstinline/z/ by the \lstinline|output| directive. Otherwise, the execution
terminates with an error message containing the content
\lstinline|'Unexpected pair'|.

The specification continues by defining
pre- and postconditions, where the precondition states the assumption
that the integer \lstinline/x/ is greater than or equal to  \lstinline/y/. The postcondition asserts that the result
\lstinline/z/ is greater than or equal to the substract of \lstinline/x/ and \lstinline/y/, which is greater or equal to zero. The symbol \lstinline/&&/ stands for logical predicate conjunction.

\subsection{More than one entrypoint}
\label{sec:more-than-one}
In SCV, we can specify several entrypoints at the same
time. This design enables us to specify and verify all entrypoints
at once and also to describe and verify relationships between
entrypoints.  Considering that, for a complete smart contract code, the input term includes the storage term before, and the output term represents the storage after the execution, there are two options of the syntax as shown in Listing~\ref {lst:multiple-entrypoint-specification-1} and Listing~\ref {lst:multiple-entrypoint-specification-2},
where the entrypoint name is written after the \lstinline/%/ symbol
and is followed by a property specification for the entrypoint. The
specification finishes with specifying entrypoint relations,
which we explain next.  The first option should be used when you want to check the output term, which may not be simply similar to the storage part of the input term. In the second option, the input term is then the pair of the parameter term and the storage term. The output term is similar to the storage term, with all variables added with the prefix \lstinline/post_/.

@@

To illustrate entrypoint relations, let us consider a smart contract
that models an online auction. The contract has two entrypoints,
\lstinline|%bid| and \lstinline|%close|, which serve as functions for
bidding and closing the contract, respectively.  The smart contract
storage is a pair containing as its first element a boolean value
\lstinline|auction_open| that indicates whether the contract is still
open for bidding or already closed. The second element is another
nested pair, where the first element stores the highest bidder's
address, and the second element is the contract owner's address. The
current highest bid is reflected in the balance of the contract.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                   Pair (highest_bidder: address)
                        (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open = true
       && Amount > pre_Balance
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open = true
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, pre_Balance, c);
  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open = true
       && Sender = contract_owner  
       && get_contract (unit, contract_owner) = 
          Some (c: contract unit)              
    post-condition := post_auction_open  = false
       && post_Balance = 0 
       && transfer_token (Unit, pre_Balance + Amount, c)

  (%create -> %bid) with (auction_open = true)  
       and (Amount > pre_Balance) 
       and get_contract (unit, highest_bidder) = 
           Some (c: contract unit)
  | (%create  -> %close) with (auction_open = true) 
       and (Sender = contract_owner)
       and get_contract (unit, contract_owner) = 
          Some (c: contract unit)
  | (%bid -> %bid) with (Amount > Balance - Amount) 
       and get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
  | (%bid -> %close) with (Sender = contract_owner) 
       and get_contract (unit, contract_owner) = 
            Some (c: contract unit)
  | not (%close -> %bid)
  | not (%close -> %close)
\end{lstlisting}

@@

Here is the first property that we want to check: if a user calls the
\lstinline/%bid/ entrypoint, when the contract is still open for
bidding and the bid (i.e., the amount transferred with the call) is
higher than the current bid (as specified in the
\lstinline/pre-condition/ field from line 8 to line 11 in Listing
\ref{lst:auction-contract-specification}) and the address of the highest bidder can be cast to a real contract represented by the variable \lstinline/c/ of type \lstinline/contract unit/ on the blockchain (lines 22-23 ), then the bid should
succeed.  In this case, the highest bidder has to be updated to the
sender (line 13 in the \lstinline/post-condition/ field), which is the
address of the caller, and the balance of the contract should be the
amount sent along with the call (line 14). Moreover, the previous
highest bidder should receive their bid, which means the auction
contract should transfer the \lstinline/pre_Balance/ to the previous
bidder, \lstinline/highest_bidder/ (line 16). 
The predicate \lstinline/transfer_token (data, amount, address)/
represents a smart contract call with three arguments: the data sent
along with the call, the amount of tokesn, and the address of the
receiver.  When the call is just a token transfer, the data field has
the value \lstinline/Unit/ of type \lstinline/unit/, indicating that
this is just a token transfer with no additional data. Another
important property is that the auction owner's address should not be
changed. That is, after calling the \lstinline/bid/ entrypoint, the
\lstinline/contract_owner/ should remain the same as before (line
14). The last property guarantees that when the contract owner closes the open contract (line 21), the contract is first closed
(line 24), and then money (\lstinline/post_Balance + Amount/) should be sent to the
owner. Hence, the final auction balance is zero (lines 25 and 26).
